---
title: "SPOCK"
author: "McCormick, Mark"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{SPOCK}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```
## Running SPOCK on sample data.

Sample data can be obtained from: https://github.com/labmccormick/SPOCKHelp .

Once you have pulled this data down and unzipped into a folder you can then test out some of SPOCK's features as reported in the paper. Below outlines how to run SPOCK using the subfolders of "RAW Data". For this example, all of the data was extracted to a folder "D:/Test/SPOCKHelp-master/RAW Data/".

1. Getting doubling times for various conditions

    To extract doubling times for one of the conditions listed in the subfolders, such  as the RAW\_WT_DT. First set the working directory to that location and copy the appropriate ladder file over: for this example copy Ladders_Lab_Strain/CENPKA.csv. Then simply run SPOCK::OGA(). This will create subfolders RAW and Results that will have a csv with the doubling times and plots of data collection.
    
```{r eval=FALSE}
##### Note this command will need to be pointed at the directory where you put the data
##### This setwd is just an example and shouldn't be cut/pasted
setwd("D:/Test/SPOCKHelp-master/RAW Data/RAW_WT_DT") 
SPOCK::OGA(laddername="CENPKA.csv")
```


2. Computing Doubling times and Chronological lifespan data for a full experimental run

    For this example the data is stored in the subfolder: RAW_OUTGROWTH_SI, so again we will start by setting our directory to that directory. For this experiment there is already an existing ladder file, so we will use this ladder file to run the code. This will generate individual results for each of the days in a file  results-(originalfilename).csv that will have the doubling times for each measured  strain for that day. Additionally this will do statistics for the doubling time for each  day and save that into the Replicate_Stats subfolder.
```{r eval=FALSE}
##### Note this command will need to be pointed at the directory where you put the data
##### This setwd is just an example and shouldn't be cut/pasted
setwd("D:/Test/SPOCKHelp-master/RAW Data/RAW_OUTGROWTH_SI/")
SPOCK::OGA(laddername = "w303aparentalladder.csv")
SPOCK::SurvivalCalc(resultspath = "Results",firstDay = 2)
```
    
3. Computing Wilcoxon rank sum statistics of CLS data

    Based on the results from the previous run (section 2 above), you should now find a subfolder Results/Survivalanalysis. In that directory there will be a file CLSstats.csv. This file will be the one you want to use to run the statistics. When you run this command you will be presented with a list of all available csv files in your current directory, make sure to type in CLSstats for the file to use. From that you will be prompted with what 2 strains you want to compare as well as a list of all available strains. From the 2 strains you pick it will give you a final p-value.
    
```{r eval=FALSE}
##### Note this command will need to be pointed at the directory where you put the data
##### This setwd is just an example and shouldn't be cut/pasted
setwd("D:/Test/SPOCKHelp-master/RAW Data/RAW_OUTGROWTH_SI/Results/Survivalanalysis")
SPOCK::wilcoxonCLS()
```
    
## Setup for your own data

1. File names for the Experiments

    SPOCK looks for files that have the following format to their names. <experiment>_Day_###.csv, where <experiment> can be any alphanumeric name for the experiment, followed by \_Day\_ (this syntax must be preserved), followed by the Day the measurements where made. Examples: EMSQ1909_Day_2.csv, EMSQ1909_Day_10.csv, EMSQ1909_Day_14.csv, w303a122020_Day_1.csv, w303a122020_Day_9.csv. The software will fail if it doesn't find the files in this general format so please take care in naming your experiment results. 

2. Annotation and expected pattern of .csv files from plate reader

    SPOCK expects input files to be from the Bioscreen C. These are .csv files of the following pattern, column 1 TIME of OD reading, columns 2-201 OD readings. Background wells should be annotated BLANK in lieu of the bioscreen naming (Well__),(or blankname="nameofblankwells" e.g. "YPD"), replicates should be named in lieu of their bioscreen column name (well__) with a unique name for each set of replicates(e.g. Well 101-105 are replicate1, rename these columns replicate1), this name must only contain alphanumeric characters. 
